Sparse Function-space Representation of Neural Networks

### 2023-09-05 workshops/icml-duality/main.tex
Polishing. Spacincing, typos, etc.

% Short section names etc
% This must be imported last!
%\usepackage{cleveref}
\usepackage[capitalise,nameinlink]{cleveref}
\crefname{section}{Sec.}{Secs.}
\crefname{algorithm}{Alg.}{Algs.}
\crefname{appendix}{App.}{Apps.}
\crefname{definition}{Def.}{Defs.}
\crefname{table}{Tab.}{Tabs.}

@@

%Need for adaptive learning methods + failures with current methods
Given an approximate inference technique, we demonstrate that NNs emit `dual' parameters which are the building blocks of a GP~\citep{csato2002sparse, adam2021dual, chang2023memory}.
In contrast to previous work that utilizes subsets of training data \citep{immer2021scalable}, our parameterization captures
information from {\em all} data points in a sparse representation, essential for scaling to deep learning data sets.
% predictive uncertainty.
Importantly, the dual parameterization can be used to 1) sparsify the GP without requiring further optimization
(e.g. variational inference) whilst capturing information from all data points, and 2) incorporate new data without retraining (in a computationally efficient manner
by conditioning on new data).
% In contrast to previous work that utilizes subsets of training data \citep{immer2021scalable}, our parameterization captures information from {\em all} data points in a sparse representation, essential for predictive uncertainty.
% Importantly, our method predicts in the same space as the original trained NN, with the benefit of avoiding the
% complexity introduced by working in weight-space and the notorious cubic complexity of vanilla GPs.
%\todo{Could make it clearer that the GP predicts in the output space while avoiding the NN parameter space}
%, a feature not present in previous approaches, while avoiding the notorious cubic complexity of vanilla GPs.
Through the dual parameterization, we establish a connection between the NN, GPs, and sparse approximations similar to sparse variational GPs~\citep{titsias2009variational}. We refer to our method as Sparse Function-space Representation (\our)---a sparse GP derived from a trained NN; see \cref{fig:teaser} for an example. %Moreover, this dual parameterization can be exploited to perform dual conditioning \citep{chang2022fantasizing}, \ie, an effective approach for conditioning on new data without needing to retrain the model (see \cref{fig:teaser}).
%As \our is in the dual parameters, we can perform dual conditioning recently shown effective in \cite{chang2022fantasizing}, that is, avoid retraining and condition new data into our model (see \cref{fig:teaser}) \todo{Effective compared to what?}.
\looseness-2

@@

We consider supervised learning with inputs $\vx_i \in \inputDomain$ and outputs $\vy_i \in \outputDomain$ (e.g. regression) or $\vy_{i} \in \{0,1\}^{C}$ (e.g. classification),
giving a data set $\dataset = \{(\vx_{i} , \vy_{i})\}_{i=1}^{N}$.
We introduce a  NN $f_\mathbf{w}: \inputDomain \to \outputDomain$ with weights $\weights \in \R^{P}$ and use a likelihood function $p(\vy \mid f_\mathbf{w}(\vx))$
to link the function values to the output $\vy$ (e.g. Categorical for classification).
For notational conciseness, we stick to scalar outputs $y_{i}$.

% \textbf{Bayesian NNs}
\textbf{BNNs}
In Bayesian deep learning, we place a prior over the weights $p(\vw)$ and aim to calculate the posterior over the weights given the data $p(\vw \mid \mathcal{D})$.
Given the weight posterior $p(\vw \mid \mathcal{D})$, we can make probabilistic predictions with,
\begin{align} \label{eq-predictions}
  p_{\text{\sc BNN}}(y \mid \vx, \mathcal{D}) = \E_{p(\vw \mid \mathcal{D})} \left[ p(y \mid f_{\vw}(\vx)) \right]
\end{align}
The posterior ${p(\vw \mid \dataset) \propto p(\vy \mid f_{\weights}(\vx)) \, p(\weights)}$ is generally not available in closed form
so we resort to approximations.
% Sampling methods that characterize the posterior with a finite set of samples in weight-space are general-purpose, but computationally heavy and impractical for downstream applications.
% In this paper, we build on the Laplace approximation

@@

\textbf{Laplace approximation}
The Laplace approximation \citet{mackayBayesian1992,daxberger2021laplace} builds upon this and approximates the weight posterior $p(\vw \mid \mathcal{D}) \approx q(\vw)= \mathcal{N}(\vw^{*} , \bm\Sigma)$
around the MAP weights ($\vw^{*}$) by setting the covariance to the Hessian of the posterior,
\begin{align} \label{eq:laplace-cov}
 \bm\Sigma = - \left[\nabla^{2}_{\vw\vw} \log p(\vw \mid \mathcal{D})  |_{\vw=\vw^{*}} \right]^{-1}.
\end{align}
Computing \cref{eq:laplace-cov} requires calculating the Hessian of the log-likelihood $\nabla^{2}_{\vw\vw} \log p(\vy \mid f_{\vw}(\vx))$ from \cref{eq-empirical-risk}.
As highlighted in \citet{immer2021improving}, computing this Hessian is often infeasible and in practice it is common to adopt the
generalized Gauss-Newton (GGN) approximation,
\begin{align}
 \nabla^{2}_{\vw\vw} \log p(\vy \mid f_{\vw}(\vx)) \approx \Jac{\vw}{\vx}^{\top} \nabla_{\vf\vf}^{2}\log(\vy\mid\vf) \Jac{\vw}{\vx}.
  \nonumber
\end{align}
where the Jacobian is given by $\Jac{\weights}{\vx} \coloneqq \left[ \nabla_\weights f_\weights(\vx)\right]^\top \in \R^{1 \times P}$.
% \textbf{Laplace-GGN converts BNN into GLM}
\citet{immer2021improving} highlighted that the GGN approximation corresponds to a local linearization of the NN,
% Recent work \citep{khan2019approximate,maddox2021fast} has converted NNs to GPs by linearizing around the NN's MAP weights (obtained from \cref{eq-empirical-risk}).
% Recent work \citep{khan2019approximate,maddox2021fast} has shown that NN (in weight space) leads to a GP, i.e. a function-space approximation.
$f_{\weights^{*}}^{\text{lin}}(\vx) = f_{\weights}(\vx) + \Jac{\weights^{*}}{\vx}(\weights-\weights^{*})$.
This suggests that predictions should be made with a generalized linear model (GLM), given by,
\begin{align} \label{eq-glm}
  p_{\text{\sc GLM}}(y_{i} \mid \vx, \mathcal{D}) &= \E_{q(\vw)} \left[ p(y_{i} \mid f_{\vw^{*}}^{\text{lin}}(\vx))\right]. %\nonumber \\
  % &\approx \frac{1}{S} \sum_{s} p(\vy \mid f_{\vw_{s}}^{\text{lin}}(\vx)), \quad \vw_{s} \sim q(\vw)
\end{align}

% \textbf{Weight space GLM is equivalent to a GP}
\textbf{Gaussian processes}
As Gaussian distributions remain tractable under linear transformations, we can convert the linear model from
weight space to function space \citep[see Ch.~2.1 in ][]{rasmussen2006gaussian}.
 % highlighted that the GGN approximation corresponds to a local linearization of the NN,
As such (and shown in \citet{immer2021improving}), the Bayesian GLM in \cref{eq-glm} has an equivalent GP formulation,
% \begin{align}
%   p_{\text{\sc GP}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vf)} \left[ p(\vy \mid \vf)\right] %\nonumber \\
%   % &\approx \frac{1}{S} \sum_{s} p(\vy \mid f_{\vw_{s}}^{\text{lin}}(\vx)), \quad \vw_{s} \sim q(\vw)
% \end{align}
% where the GP posterior is given by,
% \begin{subequations}  \label{eq-gp-pred-immer}
% \begin{align}
%   p_{\text{\sc GP}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vf)} \left[ p(\vy \mid \vf)\right] \\ %\nonumber \\
%   \myexpect_{p(f_i \mid\vy)}[f_i] &= f_{\vw^{*}}(\vx_{i}), \\
%   \mathrm{Var}_{p(f_i \mid \vy)}[f_i] &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \bm\Lambda^{-1})^{-1} \vk_{\vx i},
% \end{align}
% \end{subequations}
\begin{align} \label{eq-gp-pred-immer}
  p_{\text{\sc GP}}(y_{i} \mid \vx, \mathcal{D}) &= \E_{q(f_{i})} \left[ p(y_{i} \mid f_{i} )\right]
\end{align}
where $q(f_{i}) = \mathcal{N} \left( f_{\vw^{*}}(\vx_{i}), \sigma^{2}_{i} \right)$ is the GP's predictive posterior with variance,
% where $q(\vf) = \mathcal{N} \left( \myexpect_{p(f_i \mid\vy)}[f_i], \mathrm{Var}_{p(f_i \mid \vy)}[f_i] \right)$ is the predictive GP posterior given by,
\begin{align} \label{eq-gp-pred-immer-gp}
  % \myexpect_{p(f_i \mid\vy)}[f_i] &= f_{\vw^{*}}(\vx_{i}), \\
  \sigma^{2}_{i} &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \bm\Lambda^{-1})^{-1} \vk_{\vx i},
\end{align}
where the kernel $\kappa(\vx,\vx')=\frac{1}{\delta} \Jac{\weights^*}{\vx} \, \JacT{\weights^*}{\vx'}$ is the Neural Tangent Kernel \citep[NTK,][]{jacot2018neural}
and $f_i = f_\vw(\vx_i)$ is the function output at $\vx_i$.
The $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where
each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, $k_{ii} = \kappa(\vx_i, \vx_i)$ and $\bm\Lambda = - \nabla^2_{\vf \vf}\log p(\vy \mid \vf)$ can be interpreted as per-input noise.
% For notational conciseness, we % \cref{eq-laplace-approx-function-space},
% restrict our notatin to a single function output. The extension to multiple outputs is straightforward.

@@

Not only does the GP formulation from \citet{immer2021improving} (shown in \cref{eq-gp-pred-immer}) struggle to scale to large
data sets but it also cannot incorporate new data by conditioning on it.
This is because its posterior mean is the NN  $f_{\vw^*}(\vx)$.
Our method overcomes both of these limitations via a dual parameterization.
Importantly, our dual parameterization enables us to 1) sparsify the GP without further optimization, and 2)
incorporate new data without retraining.\looseness-1 % (in a computationally efficient manner).

@@

% \newpage
% \section{\our: Sparse function-space representation of neural networks}\label{sec:sfr}
\section{\our: Sparse function-space representation of NNs}\label{sec:sfr}
%\paragraph{GP in the dual parameters}
In this section, we present our method, named \our, which converts a trained NN into a GP.
\our is built upon a dual parameterization of the GP posterior.
Early work by \citet{csato2002sparse} showed that Expectation Propagation gives rise to a dual parameterization.
More recent work by \citet{adam2021dual,chang2023memory} showed a different dual parameterization arising from the evidence lower bound for sparse variational GPs \citep{hensman2013gaussian,titsias2009variational}.
The dual parameterization from \citet{adam2021dual,chang2023memory}, consists of parameters $\valpha$ and $\vbeta$, which gives rise to the predictive posterior,
% , we can build on the insight from \citet{csato2002sparse} to obtain a `dual' parameterization, $\valpha$ and $\vbeta$,
% The seminal work by \citet{csato2002sparse} (parameterization Lemma~1) gives a parameterization for the posterior process that can be found through the Bayesian update using the GP prior and the likelihood function. This gives a `dual' parameterization, $\valpha$ and $\vbeta$,
% %
\begin{subequations}  \label{eq:gp_pred}
\begin{align}
  \myexpect_{p(f_i \mid\vy)}[f_i] &= \vk_{\vx i}^\top \valpha, \\
  \mathrm{Var}_{p(f_i \mid \vy)}[f_i] &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i},  \nonumber
\end{align}
\end{subequations}
%
% where $f_i = f_\vw(\vx_i)$ is the function output at input $\vx_i$.
% the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $k_{ii} = \kappa(\vx_i, \vx_i)$.
\cref{eq:gp_pred} states that the first two moments of the resultant posterior process (which may not be a GP), can be parameterized via the dual
parameters $\valpha, \vbeta \in \R^{N}$,
% which are the vectors
defined as, \looseness-1
%
\begin{subequations}
\label{eq:dual_param}
\begin{align}
  \alpha_i &\coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f}\log p(y_i \mid f) |_{f=f_i}], \\
  %\quad \text{and} \quad
  \beta_i &\coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f f}\log p(y_i \mid f_i) |_{f=f_i}].
  \end{align}
\end{subequations}
%
\cref{eq:dual_param} holds for generic likelihoods and involves no approximations since the expectation is under the
exact posterior, given that the model can be expressed in a kernel formulation.
\cref{eq:gp_pred} and \cref{eq:dual_param} highlight that the approximate inference technique, usually viewed as a posterior approximation,
can alternatively be interpreted as an approximation of the expectation of loss (likelihood) gradients.

@@

\paragraph{Dual parameters from NN}
Given that we use a Laplace approximation of the NN, we remove the expectation over the posterior \citep[see Ch.3.4.1 in][for derivation]{rasmussen2006gaussian} and our dual parameters are given by,
%
\begin{subequations}
\label{eq:dual_param_laplace}
\begin{align}
  \hat{\alpha}_i &\coloneqq \nabla_{f}\log p(y_i \mid f) |_{f=f_i} , \\
  %\quad \text{and} \quad
  \hat{\beta}_i &\coloneqq - \nabla^2_{ff}\log p(y_i \mid f) |_{f=f_i}.
  \end{align}
\end{subequations}
%
Substituting \cref{eq:dual_param_laplace} into \cref{eq:gp_pred}, we obtain our GP based on the trained NN.
Making predictions with \cref{eq:gp_pred} costs $O(N^3)$, which limits its applicability on large data sets.

@@

\begin{figure*}[t]
  % \centering\scriptsize
  \centering\small
  % \setlength{\figurewidth}{.18\textwidth}
  \setlength{\figurewidth}{.27\textwidth}
  \setlength{\figureheight}{\figurewidth}
  \pgfplotsset{axis on top,scale only axis,y tick label style={rotate=90}, x tick label style={font=\Large},y tick label style={font=\LARGE},title style={yshift=-4pt,font=\LARGE}, y label style={font=\LARGE},x label style={font=\LARGE},grid=major, width=\figurewidth, height=\figureheight}
  \pgfplotsset{grid style={line width=.1pt, draw=gray!10,dashed}}
  \pgfplotsset{xlabel={$M$ as \% of $N$},ylabel style={yshift=-5pt},xlabel style={yshift=-5pt}}
  %
  \begin{minipage}[t]{.155\textwidth}
    \raggedleft
    \pgfplotsset{ylabel=NLPD}
    \input{./fig/australian.tex}
  \end{minipage}
  \hfill
  % \begin{minipage}[t]{.12\textwidth}
  %   \raggedleft
  %   \input{./fig/breast_cancer.tex}
  % \end{minipage}
  % \hfill
  % \begin{minipage}[t]{.12\textwidth}
  %   \raggedleft
  %   \input{./fig/ionosphere.tex}
  % \end{minipage}
  % \hfill
  \begin{minipage}[t]{.155\textwidth}
    \raggedleft
    \input{./fig/glass.tex}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{.155\textwidth}
    \raggedleft
    \input{./fig/vehicle.tex}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{.155\textwidth}
    \raggedleft
    \input{./fig/waveform.tex}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{.155\textwidth}
    \raggedleft
    \input{./fig/digits.tex}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{.155\textwidth}
    \raggedleft
    \input{./fig/satellite.tex}
  \end{minipage}\\[-1em]
  %
  % Legend
  \definecolor{steelblue31119180}{RGB}{31,119,180}
  \definecolor{darkorange25512714}{RGB}{255,127,14}
  \newcommand{\myline}[1]{\protect\tikz[baseline=-.5ex,line width=1.6pt]\protect\draw[draw=#1](0,0)--(1.2em,0);}
  \caption{Comparison of convergence in terms of number of inducing points $M$ in NLPD (mean\textcolor{gray}{\footnotesize$\pm$std} over 5 seeds) on UCI classification tasks: \our (\myline{steelblue31119180}) vs. GP subset (\myline{darkorange25512714}). Our \our converges fast for all cases showing clear benefits of its ability to summarize all the data onto a sparse set of inducing points. The number of inducing points $M$ are specified as a percentage of the number of training data $N$.}
  \label{fig:uci}
  %\vspace*{-6pt}
\end{figure*}

>>>>

% Short section names etc
% This must be imported last!
%\usepackage{cleveref}
\usepackage[capitalise,nameinlink]{cleveref}
\crefname{section}{Sec.}{Secs.}
\crefname{algorithm}{Alg.}{Algs.}
\crefname{appendix}{App.}{Apps.}
\crefname{definition}{Def.}{Defs.}
%\crefname{table}{Tab.}{Tabs.} % Let's leave table as is

@@

%Need for adaptive learning methods + failures with current methods
Given an approximate inference technique, we demonstrate that NNs emit `dual' parameters which are the building blocks of a GP~\citep{csato2002sparse, adam2021dual, chang2023memory}.
In contrast to previous work that utilizes subsets of training data \citep{immer2021scalable}, our parameterization captures
information from {\em all} data points in a sparse representation, essential for scaling to deep learning data sets.
% predictive uncertainty.
Importantly, the dual parameterization can be used to {\em (i)}~sparsify the GP without requiring further optimization
(\eg, variational inference) whilst capturing information from all data points, and {\em (ii)}~incorporate new data without retraining (in a computationally efficient manner
by conditioning on new data).
% In contrast to previous work that utilizes subsets of training data \citep{immer2021scalable}, our parameterization captures information from {\em all} data points in a sparse representation, essential for predictive uncertainty.
% Importantly, our method predicts in the same space as the original trained NN, with the benefit of avoiding the
% complexity introduced by working in weight-space and the notorious cubic complexity of vanilla GPs.
%\todo{Could make it clearer that the GP predicts in the output space while avoiding the NN parameter space}
%, a feature not present in previous approaches, while avoiding the notorious cubic complexity of vanilla GPs.
Through the dual parameterization, we establish a connection between the NN, GPs, and sparse approximations similar to sparse variational GPs~\citep{titsias2009variational}. We refer to our method as Sparse Function-space Representation (\our)---a sparse GP derived from a trained NN; see \cref{fig:teaser} for an example. %Moreover, this dual parameterization can be exploited to perform dual conditioning \citep{chang2022fantasizing}, \ie, an effective approach for conditioning on new data without needing to retrain the model (see \cref{fig:teaser}).
%As \our is in the dual parameters, we can perform dual conditioning recently shown effective in \cite{chang2022fantasizing}, that is, avoid retraining and condition new data into our model (see \cref{fig:teaser}) \todo{Effective compared to what?}.
\looseness-2

@@

We consider supervised learning with inputs $\vx_i \in \inputDomain$ and outputs $\vy_i \in \outputDomain$ (\eg, regression) or $\vy_{i} \in \{0,1\}^{C}$ (\eg, classification),
giving a data set $\dataset = \{(\vx_{i} , \vy_{i})\}_{i=1}^{N}$.
We introduce a  NN $f_\mathbf{w}: \inputDomain \to \outputDomain$ with weights $\weights \in \R^{P}$ and use a likelihood function $p(\vy \mid f_\mathbf{w}(\vx))$
to link the function values to the output $\vy$ (\eg, categorical for classification).
For notational conciseness, we stick to scalar outputs $y_{i}$.

% \textbf{Bayesian NNs}
\textbf{BNNs}
In Bayesian deep learning, we place a prior over the weights $p(\vw)$ and aim to calculate the posterior over the weights given the data $p(\vw \mid \mathcal{D})$.
Given the weight posterior $p(\vw \mid \mathcal{D})$, we can make probabilistic predictions with
\begin{align} \label{eq-predictions}
  p_{\text{\sc BNN}}(y \mid \vx, \mathcal{D}) = \E_{p(\vw \mid \mathcal{D})} \left[ p(y \mid f_{\vw}(\vx)) \right].
\end{align}
The posterior ${p(\vw \mid \dataset) \propto p(\vy \mid f_{\weights}(\vx)) \, p(\weights)}$ is generally not available in closed form
so we resort to approximations.
% Sampling methods that characterize the posterior with a finite set of samples in weight-space are general-purpose, but computationally heavy and impractical for downstream applications.
% In this paper, we build on the Laplace approximation

@@

\textbf{Laplace approximation}
The Laplace approximation \citet{mackayBayesian1992,daxberger2021laplace} builds upon this and approximates the weight posterior $p(\vw \mid \mathcal{D}) \approx q(\vw)= \mathcal{N}(\vw^{*} , \bm\Sigma)$
around the MAP weights ($\vw^{*}$) by setting the covariance to the Hessian of the posterior,
\begin{align} \label{eq:laplace-cov}
 \bm\Sigma = - \left[\nabla^{2}_{\vw\vw} \log p(\vw \mid \mathcal{D})  |_{\vw=\vw^{*}} \right]^{-1}.
\end{align}
Computing \cref{eq:laplace-cov} requires calculating the Hessian of the log-likelihood $\nabla^{2}_{\vw\vw} \log p(\vy \mid f_{\vw}(\vx))$ from \cref{eq-empirical-risk}.
As highlighted in \citet{immer2021improving}, computing this Hessian is often infeasible and in practice it is common to adopt the
generalized Gauss-Newton (GGN) approximation,
\begin{align}
 \nabla^{2}_{\vw\vw} \log p(\vy \mid f_{\vw}(\vx)) \approx \Jac{\vw}{\vx}^{\top} \nabla_{\vf\vf}^{2}\log(\vy\mid\vf) \Jac{\vw}{\vx}.
  \nonumber
\end{align}
where the Jacobian is given by $\Jac{\weights}{\vx} \coloneqq \left[ \nabla_\weights f_\weights(\vx)\right]^\top \in \R^{1 \times P}$.
% \textbf{Laplace-GGN converts BNN into GLM}
\citet{immer2021improving} highlighted that the GGN approximation corresponds to a local linearization of the NN,
% Recent work \citep{khan2019approximate,maddox2021fast} has converted NNs to GPs by linearizing around the NN's MAP weights (obtained from \cref{eq-empirical-risk}).
% Recent work \citep{khan2019approximate,maddox2021fast} has shown that NN (in weight space) leads to a GP, i.e. a function-space approximation.
$f_{\weights^{*}}^{\text{lin}}(\vx) = f_{\weights}(\vx) + \Jac{\weights^{*}}{\vx}(\weights-\weights^{*})$.
This suggests that predictions should be made with a generalized linear model (GLM), given by
\begin{align} \label{eq-glm}
  p_{\text{\sc GLM}}(y_{i} \mid \vx, \mathcal{D}) &= \E_{q(\vw)} \left[ p(y_{i} \mid f_{\vw^{*}}^{\text{lin}}(\vx))\right]. %\nonumber \\
  % &\approx \frac{1}{S} \sum_{s} p(\vy \mid f_{\vw_{s}}^{\text{lin}}(\vx)), \quad \vw_{s} \sim q(\vw)
\end{align}

% \textbf{Weight space GLM is equivalent to a GP}
\textbf{Gaussian processes}
As Gaussian distributions remain tractable under linear transformations, we can convert the linear model from
weight space to function space \citep[see Ch.~2.1 in ][]{rasmussen2006gaussian}.
 % highlighted that the GGN approximation corresponds to a local linearization of the NN,
As such (and shown in \citet{immer2021improving}), the Bayesian GLM in \cref{eq-glm} has an equivalent GP formulation,
% \begin{align}
%   p_{\text{\sc GP}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vf)} \left[ p(\vy \mid \vf)\right] %\nonumber \\
%   % &\approx \frac{1}{S} \sum_{s} p(\vy \mid f_{\vw_{s}}^{\text{lin}}(\vx)), \quad \vw_{s} \sim q(\vw)
% \end{align}
% where the GP posterior is given by,
% \begin{subequations}  \label{eq-gp-pred-immer}
% \begin{align}
%   p_{\text{\sc GP}}(\vy \mid \vx, \mathcal{D}) &= \E_{q(\vf)} \left[ p(\vy \mid \vf)\right] \\ %\nonumber \\
%   \myexpect_{p(f_i \mid\vy)}[f_i] &= f_{\vw^{*}}(\vx_{i}), \\
%   \mathrm{Var}_{p(f_i \mid \vy)}[f_i] &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \bm\Lambda^{-1})^{-1} \vk_{\vx i},
% \end{align}
% \end{subequations}
\begin{align} \label{eq-gp-pred-immer}
  p_{\text{\sc GP}}(y_{i} \mid \vx, \mathcal{D}) &= \E_{q(f_{i})} \left[ p(y_{i} \mid f_{i} )\right],
\end{align}
where $q(f_{i}) = \mathcal{N} \left( f_{\vw^{*}}(\vx_{i}), \sigma^{2}_{i} \right)$ is the GP's predictive posterior with variance,
% where $q(\vf) = \mathcal{N} \left( \myexpect_{p(f_i \mid\vy)}[f_i], \mathrm{Var}_{p(f_i \mid \vy)}[f_i] \right)$ is the predictive GP posterior given by,
\begin{align} \label{eq-gp-pred-immer-gp}
  % \myexpect_{p(f_i \mid\vy)}[f_i] &= f_{\vw^{*}}(\vx_{i}), \\
  \sigma^{2}_{i} &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \bm\Lambda^{-1})^{-1} \vk_{\vx i},
\end{align}
where the kernel $\kappa(\vx,\vx')=\frac{1}{\delta} \Jac{\weights^*}{\vx} \, \JacT{\weights^*}{\vx'}$ is the Neural Tangent Kernel \citep[NTK,][]{jacot2018neural}
and $f_i = f_\vw(\vx_i)$ is the function output at $\vx_i$.
The $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where
each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, $k_{ii} = \kappa(\vx_i, \vx_i)$ and $\bm\Lambda = - \nabla^2_{\vf \vf}\log p(\vy \mid \vf)$ can be interpreted as per-input noise.
% For notational conciseness, we % \cref{eq-laplace-approx-function-space},
% restrict our notatin to a single function output. The extension to multiple outputs is straightforward.

@@

Not only does the GP formulation from \citet{immer2021improving} (shown in \cref{eq-gp-pred-immer}) struggle to scale to large
data sets but it also cannot incorporate new data by conditioning on it.
This is because its posterior mean is the NN  $f_{\vw^*}(\vx)$.
Our method overcomes both of these limitations via a dual parameterization.
Importantly, our dual parameterization enables us to {\em (i)}~sparsify the GP without further optimization, and {\em (ii)}~incorporate new data without retraining.\looseness-1 % (in a computationally efficient manner).

@@

% \newpage
% \section{\our: Sparse function-space representation of neural networks}\label{sec:sfr}
\section{\our: Sparse function-space representation of NNs}\label{sec:sfr}
%\paragraph{GP in the dual parameters}
In this section, we present our method, named \our, which converts a trained NN into a GP.
\our is built upon a dual parameterization of the GP posterior.
Early work by \citet{csato2002sparse} showed that Expectation Propagation gives rise to a dual parameterization.
More recent work by \citet{adam2021dual,chang2023memory} showed a different dual parameterization arising from the evidence lower bound for sparse variational GPs \citep{hensman2013gaussian,titsias2009variational}.
The dual parameterization from \citet{adam2021dual,chang2023memory}, consists of parameters $\valpha$ and $\vbeta$, which gives rise to the predictive posterior,
% , we can build on the insight from \citet{csato2002sparse} to obtain a `dual' parameterization, $\valpha$ and $\vbeta$,
% The seminal work by \citet{csato2002sparse} (parameterization Lemma~1) gives a parameterization for the posterior process that can be found through the Bayesian update using the GP prior and the likelihood function. This gives a `dual' parameterization, $\valpha$ and $\vbeta$,
% %
\begin{subequations}  \label{eq:gp_pred}
\begin{align}
  \myexpect_{p(f_i \mid\vy)}[f_i] &= \vk_{\vx i}^\top \valpha, \\
  \mathrm{Var}_{p(f_i \mid \vy)}[f_i] &= k_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i}.  \nonumber
\end{align}
\end{subequations}
%
% where $f_i = f_\vw(\vx_i)$ is the function output at input $\vx_i$.
% the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $k_{ii} = \kappa(\vx_i, \vx_i)$.
\cref{eq:gp_pred} states that the first two moments of the resultant posterior process (which may not be a GP), can be parameterized via the dual
parameters $\valpha, \vbeta \in \R^{N}$,
% which are the vectors
defined as \looseness-1
%
\begin{subequations}
\label{eq:dual_param}
\begin{align}
  \alpha_i &\coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f}\log p(y_i \mid f) |_{f=f_i}], \\
  %\quad \text{and} \quad
  \beta_i &\coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f f}\log p(y_i \mid f_i) |_{f=f_i}].
  \end{align}
\end{subequations}
%
\cref{eq:dual_param} holds for generic likelihoods and involves no approximations since the expectation is under the
exact posterior, given that the model can be expressed in a kernel formulation.
\cref{eq:gp_pred} and \cref{eq:dual_param} highlight that the approximate inference technique, usually viewed as a posterior approximation,
can alternatively be interpreted as an approximation of the expectation of loss (likelihood) gradients.

@@

\paragraph{Dual parameters from NN}
Given that we use a Laplace approximation of the NN, we remove the expectation over the posterior \citep[see Ch.~3.4.1 in][for derivation]{rasmussen2006gaussian} and our dual parameters are given by
%
\begin{subequations}
\label{eq:dual_param_laplace}
\begin{align}
  \hat{\alpha}_i &\coloneqq \nabla_{f}\log p(y_i \mid f) |_{f=f_i} , \\
  %\quad \text{and} \quad
  \hat{\beta}_i &\coloneqq - \nabla^2_{ff}\log p(y_i \mid f) |_{f=f_i}.
  \end{align}
\end{subequations}
%
Substituting \cref{eq:dual_param_laplace} into \cref{eq:gp_pred}, we obtain our GP based on the trained NN.
Making predictions with \cref{eq:gp_pred} costs $O(N^3)$, which limits its applicability on large data sets.

@@

\begin{figure*}[t]
  % \centering\scriptsize
  \centering\small
  % \setlength{\figurewidth}{.18\textwidth}
  \setlength{\figurewidth}{.27\textwidth}
  \setlength{\figureheight}{\figurewidth}
  \pgfplotsset{axis on top,scale only axis,y tick label style={rotate=90}, x tick label style={font=\Large},y tick label style={font=\Large},title style={yshift=-4pt,font=\LARGE}, y label style={font=\Large},x label style={font=\LARGE},grid=major, width=\figurewidth, height=\figureheight}
  \pgfplotsset{grid style={line width=.1pt, draw=gray!10,dashed}}
  \pgfplotsset{xlabel={$M$ as \% of $N$},ylabel style={yshift=-5pt},xlabel style={yshift=-5pt}}
  %
  \begin{minipage}[t]{.175\textwidth}
    \raggedleft
    \pgfplotsset{ylabel=NLPD}
    \input{./fig/australian.tex}
  \end{minipage}
  \hfill
  % \begin{minipage}[t]{.12\textwidth}
  %   \raggedleft
  %   \input{./fig/breast_cancer.tex}
  % \end{minipage}
  % \hfill
  % \begin{minipage}[t]{.12\textwidth}
  %   \raggedleft
  %   \input{./fig/ionosphere.tex}
  % \end{minipage}
  % \hfill
  \begin{minipage}[t]{.155\textwidth}
    \raggedleft
    \input{./fig/glass.tex}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{.155\textwidth}
    \raggedleft
    \input{./fig/vehicle.tex}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{.155\textwidth}
    \raggedleft
    \input{./fig/waveform.tex}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{.155\textwidth}
    \raggedleft
    \input{./fig/digits.tex}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{.155\textwidth}
    \raggedleft
    \input{./fig/satellite.tex}
  \end{minipage}\\[-1em]
  %
  % Legend
  \definecolor{steelblue31119180}{RGB}{31,119,180}
  \definecolor{darkorange25512714}{RGB}{255,127,14}
  \newcommand{\myline}[1]{\protect\tikz[baseline=-.5ex,line width=1.6pt]\protect\draw[draw=#1](0,0)--(1.2em,0);}
  \caption{Comparison of convergence in terms of number of inducing points $M$ in NLPD (mean\textcolor{gray}{\footnotesize$\pm$std} over 5 seeds) on UCI classification tasks: \our (\myline{steelblue31119180}) vs. GP subset (\myline{darkorange25512714}). Our \our converges fast for all cases showing clear benefits of its ability to summarize all the data onto a sparse set of inducing points. The number of inducing points $M$ are specified as a percentage of the number of training data $N$.}
  \label{fig:uci}
  %\vspace*{-6pt}
\end{figure*}
